Rationality, Evolution \& the Environment

### 2015-11-04 paper/paper.tex
Paolo's edits Section 3; before drastic rewrite

If an agent does not have a concrete probabilistic belief (see also
Section~\ref{sec:impr-prob-beli} below), other rules of choice are available. Perhaps the most
well-known, the \emph{maximin rule}, would tell the agent to choose act $\act^*$ only if it
maximizes payoffs for all worst case scenarios. For a strategic game with 2 players this can be
formalized as
$\act^* \in \argmax_{\myact{i} \in \Acts_i} \min_{\myact{j} \in \Acts_j} \pi_i(\myact{i},
\myact{j})$.
For the example in Figure~\ref{coordgame1}, the maximim rule looks at the minima in each row,
asking: what is the worst possible out come for each row act? It is $0$ in either
case. Consequently, the maximin rule would not favor either act; players choosing by the
maximin rule would be indifferent between acts $I$ and $II$.

@@

Regret minimization is intimately related to the maximin rule. Let the \emph{negative regret
  transform} of a strategic game $G$ be the game $G'$ derived from $G$ by replacing the payoff
function $\pi$ of $G$ with $\pi'$ such that $\pi'_i(a_i, a_{-i}) = - reg(a_i,a_{-1})$, i.e.,
the payoff of all players is the pairwise negative regret (under the payoffs from $G$). For
example, the negative regret transform of the game in Figure~\ref{coordgame1} would have the
following payoff matrix, i.e., the negatives of the pairwise regrets given in
Figure~\ref{coordgame1reg}:

@@

\noindent The acts selected by the maximin rule for the negative regret transpose are exactly
the acts selected by regret minimization for the original game.

@@

In sum, we have so far considered four different types of choice rules, each differing along a
preference and a belief dimension. As for preference representations, agents could either
target objective payoffs, or consider non-veridical subjective regret transformations. As for
beliefs, agents could either hold probabilistic belief or not. In the former case, we would
assume that they maximize expectations under whatever preference representation they hold; in
the latter, we would assume that they play maximin based on their preference
representations. There are doubtlessly many more possible preference and belief types, but
these are the main characters in the plot to follow.

Before we turn to evolutionary competition between choice mechanisms, we should justify our
focus on the two choice rules. In the following we argue that both the minimax rule, as well as
the Laplace rule can be viewed as the same choice rule under different representations of
uncertainty.

@@

Consider the following problem. I have, on the desk in front of me, a bag containing either red or black marbles. There is no further information about the distribution of the two colors. What is then the probability that I will draw a red marble? As in Walley91-96, we are not talking about physical probability, but rather about epistemic probability, i.e., the subjective probability that an agent attributes to an event.
In the literature there are two traditional ways of answering this kind of questions. According to standard Bayesianism, one should always be able to determine a precise betting rate (that immediately translates into a precise probability distribution), which is fair, in the sense that the DM is indifferent between betting on or against red at the fair rate, and would bet on red at any rate that is more favourable than the fair rate and against red at any rate that is less favourable. In this case, a strict Bayesian, in the absence of any further information, would probably rely on the principle of insufficient reason and answer that the DM should have a uniform distribution that considers all the outcomes equipossible. \\
The second answer involves imprecise probabilities and is related to the literature about uncertainty, or ambiguity. Many authors argued in favor of this kind of approach [e.g., Walley91-96, Levi74, GardenforsANDSahlin82]. Since it is consistent with the information available that all the marbles are red, or that no marble is red, the resulting model should take it into account and specify a lower probability of red $\underline{P}(R)$ and an upper probability of red $\overline{P}(R)$ that define an interval of possible probabilistic beliefs [$\underline{P}(R), \overline{P}(R)$]. Given the information available in this case we would have $\underline{P}(R)=0$ and $\overline{P}(R)=1$. This approach appears particularly relevant and appropriate in a game theoretical context. Indeed, in a recent paper [Battigalliet al.15] Battigalli et al. write
\begin{quote}
Such uncertainty is inherent in situations of strategic interaction. This is quite obvious when such situations have been faced only a few times.
\end{quote}
In decision theory, the imprecise probability model is in line with most of the representation results about uncertainty [GilboaSchmeidler89,KMM06], and seems justified by empirical observations too (see below). Walley91-96 is a seminal work in the area, that also introduces learning procedures to get useful inferences and narrow down the probability set [$\underline{P}(R), \overline{P}(R)$] after observing data. Since we want to study the evolution of meta-types, the present work does not address the probem of drawing inferences from observations, and for the moment we do not endow the players with any learning procedure.\\
In line with traditional evolutionary game theory, we model a population of rather unsophisticated and myopic agents. %that possibly differ from each other with respect to two components: the player type and the epistemic type. The players 
The players encode both a psychological subjective representation of the interaction (player type) and a conceptualization of uncertainty (epistemic type), but they are too simple to have any learning or belief updating procedure. From this point of view, we can take the learning process away from natural selection and focus on player types and epistemic types as the target of evolutionary pressure.

@@

@@

\noindent To show
the result we will use the following partition of $\mathcal{G}$.
\begin{enumerate}
\item Coordination games $\mathcal{G}^{C}$: $a>c$ and $d>b$;
\item Anti-coordination games $\mathcal{G}^{A}$: $a<c$ and $d<b$;
\item Strong dominance games $\mathcal{G}^{S}$: aut $(a>c$ and $b>d)$
aut $(a<c$ and $b<d)$;
\item Weak dominance games $\mathcal{G}^{W}$: aut $a=c$ aut $b=d$;
\item Indifferent games $\mathcal{G}^{I}$: $a=c$ and $b=d$.
\end{enumerate}
Before proving the lemmas, it is convenient to fix some notation.
We denote by $G^{C}$ a coordination game in $\mathcal{G}^{C}$ with
payoffs $a^{G^{C}}$, $b^{G^{C}}$, $c^{G^{C}}$, and $d^{G^{C}}$;
similarly, for $G^{A}$, $G^{S}$, $G^{W}$, and $G^{I}$. Let us
call $I_{G^{C}}^{RM}$ the event that a RM player plays action $I$
in the game $G^{C}$; and similarly for action $II$, for player MM,
and for games $G^{A}$, $G^{S}$, $G^{W}$ and $G^{I}$. Moreover,
let us recall the definition of evolutionarily stable strategy and
neutrally stable strategy.

@@

Now let us split into $a>d$ and $a<d$, and consider $a>d$ first.
Notice that, by Lemma 7, the case $a=d$ is irrelevant in order to
discriminate between RM and MM. If $a>d$, by Lemma 7 we can eliminate
the cases where RM plays $II$ and MM plays $I$:

>>>>

If an agent does not have a precise probabilistic belief (see also
Section~\ref{sec:impr-prob-beli} below), other rules of choice are available. Perhaps the most
well-known, the \emph{maximin rule}, would tell the agent to choose act $\act^*$ only if it
maximizes payoffs for all worst case scenarios. For a strategic game with 2 players this can be
formalized as
$\act^* \in \argmax_{\myact{i} \in \Acts_i} \min_{\myact{j} \in \Acts_j} \pi_i(\myact{i},
\myact{j})$.
For the example in Figure~\ref{coordgame1}, the maximim rule looks at the minima in each row,
asking: what is the worst possible out come for each row act? It is $0$ in either
case. Consequently, the maximin rule would not favor either act; players choosing by the
maximin rule would be indifferent between acts $I$ and $II$.

@@

Regret minimization is intimately related to the maximin rule. Let the \emph{negative regret
  transformation} of a strategic game $G$ be the game $G'$ derived from $G$ by replacing the payoff
function $\pi$ of $G$ with $\pi'$ such that $\pi'_i(a_i, a_{-i}) = - reg(a_i,a_{-1})$, i.e.,
the payoff of all players is the pairwise negative regret (under the payoffs from $G$). For
example, the negative regret transform of the game in Figure~\ref{coordgame1} would have the
following payoff matrix, i.e., the negatives of the pairwise regrets given in
Figure~\ref{coordgame1reg}:

@@

\noindent The acts selected by the maximin rule for the negative regret transformation are exactly
the acts selected by regret minimization for the original game.

@@

In sum, we have so far considered four different alternatives, each differing along a
preference and a belief dimension. As for preference representations, agents could either
target objective payoffs, or consider non-veridical subjective regret transformations. Given a game $G$, we will denote by $ \tau^{reg}(G)$ the regret representation (corresponding to the negative regret transformation) of the game, and $\tau^{\pi}(G)$ the objective representation (corresponding to the identity transformation) of the game. \\
As for
beliefs, agents could either hold probabilistic belief or not. In the former case, we would
assume that they maximize expectations under whatever preference representation they hold; in
the latter, we would assume that they play maximin based on their preference
representations. There are doubtlessly many more possible preference and belief types, but
these are the main characters in the plot to follow.

Before we turn to evolutionary competition between choice principles, we should justify our
focus on the two choice rules. In the following we argue that both the minimax rule, as well as the maximization of expected utility 
(and the Laplace rule as a special case) can be viewed as the same choice rule under different representations of
uncertainty.

@@

Consider the following problem. I have, on the desk in front of me, a bag containing either red or black marbles. There is no further information about the distribution of the two colors. What is then the probability that I will draw a red marble? As in Walley91-96, we are not talking about physical probability, but rather about epistemic probability, i.e., the subjective probability that an agent attributes to an event.
In the literature there are two traditional ways of answering this kind of questions. According to standard Bayesianism, one should always be able to determine a precise betting rate (that immediately translates into a precise probability distribution), which is fair, in the sense that the DM is indifferent between betting on or against red at the fair rate, and would bet on red at any rate that is more favourable than the fair rate and against red at any rate that is less favourable. In this case, a strict Bayesian, in the absence of any further information, would probably rely on the principle of insufficient reason and answer that the DM should have a uniform distribution that considers all the outcomes equipossible. \\
The second answer involves imprecise probabilities and is related to the literature about unmeasurable uncertainty. Many authors argued in favor of this kind of approach [e.g., Walley91-96, Levi74, GardenforsANDSahlin82]. Since it is consistent with the information available that all the marbles are red, or that no marble is red, the resulting model should take it into account and specify a lower probability of red $\underline{P}(R)$ and an upper probability of red $\overline{P}(R)$ that define an interval of possible probabilistic beliefs [$\underline{P}(R), \overline{P}(R)$]. Given the information available in this case we would have $\underline{P}(R)=0$ and $\overline{P}(R)=1$. This approach appears particularly relevant and appropriate in a game theoretical context. Indeed, in a recent paper [Battigalliet al.15] Battigalli et al. write

\begin{quote}
Such uncertainty is inherent in situations of strategic interaction. This is quite obvious when such situations have been faced only a few times.
\end{quote}

In decision theory, the imprecise probability model is in line with most of the representation results about uncertainty [GilboaSchmeidler89,KMM06], and seems justified by empirical observations too.

A prominent example is Ellsberg's paradox. A bag contains 90 marbles, 30 are red and the remaining 60 are either black or yellow. A marble is to be drawn and DM can win 100€ if she guesses the color of the drawn marble. The DM is offered two different bets. In the first one she can choose to bet either on red (R) or on yellow (Y), while in the second she can bet either on red or black (RB) or on yellow or black (YB), as shown in Table \ref{Ellsberg}.

\definecolor{yellow}{RGB}{255,188,1}

\begin{table}[h]
\centering
\begin{tabular}{@{}llll@{}}
\cmidrule(l){2-4}
\multicolumn{1}{c}{} & {\color{red}R}   & {\color{yellow}Y}   & B   \\ \cmidrule(l){2-4} 
$f_{{\color{red}R}}$              & 100 & 0   & 0   \\
$f_{{\color{yellow}Y}}$              & 0   & 100 & 0   \\
$f_{{\color{red}R}B}$            & 100 & 0   & 100 \\
$f_{{\color{yellow}Y}B}$             & 0   & 100 & 100 \\ \bottomrule
\end{tabular}
\caption{Ellsberg's paradox}
\label{Ellsberg}
\end{table}

The pattern consistently observed in experimets is: $f_{{\color{red}R}} \succ f_{{\color{yellow}Y}}$ and $f_{{\color{red}R}B} \prec f_{{\color{yellow}Y}B}$. This is clearly incompatible with any precise probabilistic belief about the proportion of black and yellow marbles. This kind of behavior has been
famously axiomatized by Gilboa and Schmeidler in a paper from 1989. The choice pattern is normally explained in terms of aversion to uncertainty, and the standard representation is given by means of probability sets together with maxmin rule. We can think of the DM in Ellsberg's example as maximinimizing expected utility over the probability set $ [\underline{P}(R)=\frac{1}{3},\underline{P}(Y)=0,\underline{P}(B)=0,\overline{P}(R)=\frac{1}{3},\overline{P}(Y)=\frac{2}{3},\overline{P}(B)=\frac{2}{3}]  $. This amounts to calculating for any probability distribution contained in the probability set the expected utility of any available action, and then to choosing the action that guarantees the highest minimal expected utility. The following observation is then trivial.

\begin{fact} \label{fact:singleton probability set}
Whenever the probability set is a singleton, maximinization of
expected utility coincides with maximization of expected utility.
\end{fact}

A huge body of experimental literature strongly suggests that agents are generally uncertainty
averse [REF]. In line with empirical data,
we assume that players in the population are uncertainty averse and conform to maxmin expected
utility. The resulting behavior is then produced by maximinimizing the subjective utility
given by the player type over the set of probabilities given by the belief state, or epistemic type, of the agent.

In full generality, we think of an epistemic type as a general disposition to form beliefs about the co-player’s behavior.    Players in the population obviously face uncertainty about the composition of the population that they are part of, and consequently about the type of co-player that they are randomly paired with at each round and about the co-player's actions. We represent this uncertainty in two different ways that correspond to the two approaches discussed above: a Bayesian uniform distribution and an imprecise probability set. This amounts to having two different epistemic types, $ \overline{\mu}$ and $ \Delta(A_{-i}) \rbrace $, where $\overline{\mu}$ stands for the uniform distribution and $\Delta(A_{-i})$ is the probability set, defined as follows:

\item $\overline{\mu} \in \Delta(A_{-i})$ is the probability measure over the co-player's actions such that: $\forall a_{-i} \in A_{-i}$, $\overline{\mu}(a_{-i})= \frac{1}{|A_{-i}|}$.

To sum up, here we consider the choice of an action as depending on two features: the subjective preference and the belief state. Definition \ref{def:choice rule} makes this point clear.

\begin{definition} \label{def:choice rule}

Given a set of subjective preferences $\Theta$ and a set of epistemic types $E$, we define \textit{choice rule} a function $\kappa: \Theta \times E  \rightarrow A_i$ that associates an action choice to a pair $(\theta, e) \in \Theta \times E$.

According to what we argued in this section, we will assume that the choice rule that turns subjective preferences and beliefs into actions is maxmin expected utility.

\begin{example}
  Consider the game in Figure \ref{coordgame1} again. When a player characterized
  by\footnote{To ease notation in the following we will refer to a player directly as a pair
    $(player \text{ } type \text{ } \tau,\text{ } epistemic\text{ } type \text{ } e)$.} player
  type $ \tau^{reg} $ and epistemic type $ \Delta(A_{-i}) $ faces the game in Figure
  \ref{coordgame1}, her player type gives rise to the representation $\tau^{reg}(G^C)$ in
  Figure \ref{coordgame1reg}. Then, maximinimizing expected utility over the probability
  simplex $ \Delta(A_{-i}) $ given the game in Figure \ref{coordgame1reg} corresponds to
  playing the standard maxmin strategy [REF] of that game.\footnote{Notice that this is the
    case for any game: given a game $G$, maximinimizing expected utility over the probability
    simplex $ \Delta(A_{-i}) $ amounts to playing the standard maxmin strategy of $G$.} This,
  in turn, corresponds to the regret minimizing action of the original game $G^C$ of Figure
  \ref{coordgame1}, namely action $II$, as we have seen in Section \ref{sec:regreTheory}. A
  player $(\tau^{reg},\overline{\mu})$ instead has subjective representation $\tau^{reg}(G^C)$
  of Figure \ref{coordgame1reg} and a flat probability measure
  $\overline{\mu}=(\frac{1}{2}I,\frac{1}{2}II)$ over the co-player's actions. Consequently,
  such a player would play action $II$, i.e., the action that maximinimizes expected utility
  given the flat probability measure $\overline{\mu}$.\footnote{As we already noticed,
    maximinimizing expected utility given a probability measure $\mu$ is equivalent to
    maximizing expected utility given $\mu$.}  Similarly, a player
  $(\tau^{\pi},\Delta(A_{-i}))$ will have the subjective representation coinciding with the
  objective fitness game, i.e., $ \tau^{\pi}(G^C)= G^C $. Then, both actions $I$ and $II$
  maximinimize expected utility over the probability simplex $ \Delta(A_{-i}) $. Hence, a
  player $(\tau^{\pi},\Delta(A_{-i}))$ will be indifferent between $I$ and $II$, and will
  randomize with equal probability $(\frac{1}{2}I,\frac{1}{2}II)$. Finally, a player
  $(\tau^{\pi},\overline{\mu})$ maximinimizes the objective fitness game $G^C$ of Figure
  \ref{coordgame1} with flat probability $\overline{\mu}$ over the co-player's actions, i.e.,
  maximizes objective utility given $\overline{\mu}$. In game $G^C$ this yields action $II$.

\begin{figure}
\begin{center}%
\begin{tabular}{|c|c|c|}
\hline 
 & I & II\tabularnewline
\hline 
\hline 
I & 0;0 & -2;-1\tabularnewline
\hline 
II & -1;-2 & 0;0\tabularnewline
\hline 
\end{tabular}\end{center}

\protect\caption{Regret representation $\tau^{reg}(G^C)$}
\label{coordgame1reg}
\end{figure}

We are now in the position of deriving behavior directly from the pair $(\tau, e)$. Definitions \ref{def:utility-belief} and \ref{def:maxminSEU} simply formalize the notions of subjective expected utility and maxmin subjective expective utility, respectively.

\begin{definition} \label{def:utility-belief}

Given a game $G$, an action $a_i \in A_i$, and a belief $\mu \in \lbrace \mu' \in \mathbb{R}^{A_{-i}}: \sum_{a_{-i} \in A_{-i}} \mu'(a_{-i}) = 1 \rbrace$, we define $\tau(G)(a_i, \mu):= \sum_{a_{-i} \in A_{-i}} \tau(G)(a_i, a_{-i}) \cdot \mu(a_{-i})$.

\begin{definition} \label{def:maxminSEU}

Given a game $G$ and a player $(\tau, e)$, we define $\text{maxmin}(\tau(G), e):= \lbrace a'_i \in A_i : a_i \in \text{max}_{a_i \in A_i} \text{ min}_{\mu \in e} \text{ } \tau(G)(a_i, \mu) \rbrace$.

\noindent Definition \ref{def:expected fitness} gives the expected evolutionary fitness of player $(\tau, e)$ against player $(\tau', e')$ in game $G$.\footnote{As already specified in the previous example, we assume that, whenever the set $\text{maxmin}(\tau(G), e)$ is not a singleton, the player $(\tau, e)$ will mix with equal probability among the actions in $\text{maxmin}(\tau(G), e)$.}

\begin{definition} \label{def:expected fitness}

Given a game $G \in \mathcal{G}$, a player $(\tau, e)$, and a player $(\tau', e')$, we define $\pi^G((\tau, e),(\tau', e')):= \pi^G(\text{maxmin}(\tau(G), e),\text{maxmin}(\tau'(G), e'))$.

%\todo[inline]{should we explain here, or very soon, what the resulting types really are: regret minimizers, maximin players \dots}

Many other alternatives different from maxmin could produce behavior: choice rules like minmax, maxmax, and minmin would similarly generate behavior in case of a lack of precise probabilistic belief. Here we postulate that players use maxmin for two main reasons that we will expand in the next section. First, among other possible choice rules maxmin is not only intuitively appealing but also corroborated from an empirical point of view. Secondly, as it will be clear in the following, our main interest is the evolution of different patterns of subjective conceptualizations of games (player types), and we want to show that the veridical representation (objective type) is not necessarily the fittest and may not be evolutionarily stable. Assuming that all players use a fixed rule (maxmin) independent of their player type allows us to factor the choice rule out of the selection process and to focus on player types.

Before concluding this section, it is worth stressing that player $(\tau^{\pi}, \Delta(A_{-i}))$ turns out to be the objective maximinimizer, player $(\tau^{reg}, \Delta(A_{-i}))$ is the regret minimizer, player $(\tau^{\pi}, \overline{\mu})$ is the objective expected utility maximizer, and player $(\tau^{reg}, \overline{\mu})$ is the expected regret minimizer. Moreover, the following fact holds.

\iffalse

\begin{fact} \label{fact:maxEU-minReg}

For any probability measure $\mu$, maximization of expected utility and minimization of expected regret are behaviorally equivalent.

\end{fact}

\noindent As we have seen in the previous example, this is not the case in the presence of (radical) uncertainty, in the sense of representation of belief state in terms of a set of probability, as in $ \Delta(A_{-i}) $.
\fi

\begin{fact} \label{fact:equivalence2x2}

In the class of $2 \times 2$ symmetric games, players $(\tau^{reg}, \Delta(A_{-i}))$, $(\tau^{\pi}, \overline{\mu})$, and $(\tau^{reg},\overline{\mu})$ are behaviorally equivalent.

\end{fact}

\noindent Two remarks. Firstly, Fact \ref{fact:equivalence2x2} does not hold for $n \times n$
symmetric games in general. Secondly, we treat behaviorally equivalent types as distinct,
because we assume that what is selected by biological evolution is not choice behavior as such,
but the mechanism which generates behavior. This matters in particular if we assume that
mutations affect different components of the behavior generating mechanisms (see below).

\iffalse

\subsection{Epistemic types}
\label{sec:epistemic-types}

\todo[inline]{I think we really need a better (and earlier) justification for these epistemic
  types; we could appeal to information available to agents; naturalness of situations in which
  we cannot attach numbers to our beliefs (life on Mars again) \dots}

Consider the following problem. I have, on the desk in front of me, a bag containing either red or black marbles. There is no further information about the distribution of the two colors. What is then the probability that I will draw a red marble? As in Walley91-96, we are not talking about physical probability, but rather about epistemic probability, i.e., the subjective probability that an agent attributes to an event.
In the literature there are two traditional ways of answering this kind of questions. According to standard Bayesianism, one should always be able to determine a precise betting rate (that immediately translates into a precise probability distribution), which is fair, in the sense that the DM is indifferent between betting on or against red at the fair rate, and would bet on red at any rate that is more favourable than the fair rate and against red at any rate that is less favourable. In this case, a strict Bayesian, in the absence of any further information, would probably rely on the principle of insufficient reason and answer that the DM should have a uniform distribution that considers all the outcomes equipossible. \\
The second answer involves imprecise probabilities and is related to the literature about uncertainty, or ambiguity. Many authors argued in favor of this kind of approach [e.g., Walley91-96, Levi74, GardenforsANDSahlin82]. Since it is consistent with the information available that all the marbles are red, or that no marble is red, the resulting model should take it into account and specify a lower probability of red $\underline{P}(R)$ and an upper probability of red $\overline{P}(R)$ that define an interval of possible probabilistic beliefs [$\underline{P}(R), \overline{P}(R)$]. Given the information available in this case we would have $\underline{P}(R)=0$ and $\overline{P}(R)=1$. This approach appears particularly relevant and appropriate in a game theoretical context. Indeed, in a recent paper [Battigalliet al.15] Battigalli et al. write

\begin{quote}
Such uncertainty is inherent in situations of strategic interaction. This is quite obvious when such situations have been faced only a few times.
\end{quote}

In decision theory, the imprecise probability model is in line with most of the representation results about uncertainty [GilboaSchmeidler89,KMM06], and seems justified by empirical observations too (see below). Walley91-96 is a seminal work in the area, that also introduces learning procedures to get useful inferences and narrow down the probability set [$\underline{P}(R), \overline{P}(R)$] after observing data. Since we want to study the evolution of meta-types, the present work does not address the probem of drawing inferences from observations, and for the moment we do not endow the players with any learning procedure.\\
In line with traditional evolutionary game theory, we model a population of rather unsophisticated and myopic agents. %that possibly differ from each other with respect to two components: the player type and the epistemic type. The players 
The players encode both a psychological subjective representation of the interaction (player type) and a conceptualization of uncertainty (epistemic type), but they are too simple to have any learning or belief updating procedure. From this point of view, we can take the learning process away from natural selection and focus on player types and epistemic types as the target of evolutionary pressure.

@@

\subsection{Uncertainty aversion and maxmin rule}

\todo[inline]{this paragraph seems to presuppose familiarity with maxi-minimization (can we
  write this with a hyphen, any chance?); we should rewrite this or introduce the notion immediately;}

An important part of the latest literature, both experimental and theoretical, in decision theory deals with the issue of uncertainty, also called ambiguity. Technically, there is uncertainty about an event when no precise objective probability distribution about the possible outcomes is given. A prominent example is the above mentioned Ellsberg's paradox. A bag contains 90 marbles, 30 are red and the remaining 60 are either black or yellow. A marble is to be drawn and DM can win 100€ if she guesses the color of the drawn marble. The DM is offered two different bets. In the first one she can choose to bet either on red (R) or on yellow (Y), while in the second she can bet either on red or black (RB) or on yellow or black (YB), as shown in Table \ref{Ellsberg}.

\definecolor{yellow}{RGB}{255,188,1}

\begin{table}[h]
\centering
\begin{tabular}{@{}llll@{}}
\cmidrule(l){2-4}
\multicolumn{1}{c}{} & {\color{red}R}   & {\color{yellow}Y}   & B   \\ \cmidrule(l){2-4} 
$f_{{\color{red}R}}$              & 100 & 0   & 0   \\
$f_{{\color{yellow}Y}}$              & 0   & 100 & 0   \\
$f_{{\color{red}R}B}$            & 100 & 0   & 100 \\
$f_{{\color{yellow}Y}B}$             & 0   & 100 & 100 \\ \bottomrule
\end{tabular}
\caption{Ellsberg's paradox}
\label{Ellsberg}
\end{table}

The pattern consistently observed in experimets is: $f_{{\color{red}R}} \succ f_{{\color{yellow}Y}}$ and $f_{{\color{red}R}B} \prec f_{{\color{yellow}Y}B}$. This is clearly incompatible with any precise probabilistic belief about the proportion of black and yellow marbles. This kind of behavior has been
famously axiomatized by Gilboa and Schmeidler in a paper from 1989. The choice pattern is normally explained in terms of aversion to uncertainty, and the standard representation is given by means of probability sets together with maxmin rule. We can think of the DM in Ellsberg's example as maximinimizing expected utility over the probability set $ [\underline{P}(R)=\frac{1}{3},\underline{P}(Y)=0,\underline{P}(B)=0,\overline{P}(R)=\frac{1}{3},\overline{P}(Y)=\frac{2}{3},\overline{P}(B)=\frac{2}{3}]  $. This amounts to calculating for any probability distribution contained in the probability set the expected utility of any available action, and then to choosing the action that guarantees the highest minimal expected utility. Whenever the set of priors is a singleton, maximinization of
expected utility coincides with maximization of expected utility.\\
%Given a player type and an epistemic type, there are different ways to produce a behavior. 
A huge body of experimental literature strongly suggests that agents are generally uncertainty
(or ambiguity) averse [REF]. In line with empirical data,
we assume that players in the population are uncertainty averse and conform to maxmin expected
utility. The resulting behavior is then produced by maximinimizing the subjective utility
given by the player type over the set of priors given by the epistemic type.

  \todo[inline]{I would like to suggest to move Figure 2 next to Figure 1; it is already useful
  for example 1 and could be referred to there already; I would suggest to use subfigures to
  put both next to each other}

\begin{example}
  Consider the game in Figure \ref{coordgame1} again. When a player characterized
  by\footnote{To ease notation in the following we will refer to a player directly as a pair
    $(player \text{ } type \text{ } \tau,\text{ } epistemic\text{ } type \text{ } e)$.} player
  type $ \tau^{reg} $ and epistemic type $ \Delta(A_{-i}) $ faces the game in Figure
  \ref{coordgame1}, her player type gives rise to the representation $\tau^{reg}(G^C)$ in
  Figure \ref{coordgame1reg}. Then, maximinimizing expected utility over the probability
  simplex $ \Delta(A_{-i}) $ given the game in Figure \ref{coordgame1reg} corresponds to
  playing the standard maxmin strategy [REF] of that game.\footnote{Notice that this is the
    case for any game: given a game $G$, maximinimizing expected utility over the probability
    simplex $ \Delta(A_{-i}) $ amounts to playing the standard maxmin strategy of $G$.} This,
  in turn, corresponds to the regret minimizing action of the original game $G^C$ of Figure
  \ref{coordgame1}, namely action $II$, as we have seen in Section \ref{sec:regreTheory}. A
  player $(\tau^{reg},\overline{\mu})$ instead has subjective representation $\tau^{reg}(G^C)$
  of Figure \ref{coordgame1reg} and a flat probability measure
  $\overline{\mu}=(\frac{1}{2}I,\frac{1}{2}II)$ over the co-player's actions. Consequently,
  such a player would play action $II$, i.e., the action that maximinimizes expected utility
  given the flat probability measure $\overline{\mu}$.\footnote{As we already noticed,
    maximinimizing expected utility given a probability measure $\mu$ is equivalent to
    maximizing expected utility given $\mu$.}  Similarly, a player
  $(\tau^{\pi},\Delta(A_{-i}))$ will have the subjective representation coinciding with the
  objective fitness game, i.e., $ \tau^{\pi}(G^C)= G^C $. Then, both actions $I$ and $II$
  maximinimize expected utility over the probability simplex $ \Delta(A_{-i}) $. Hence, a
  player $(\tau^{\pi},\Delta(A_{-i}))$ will be indifferent between $I$ and $II$, and will
  randomize with equal probability $(\frac{1}{2}I,\frac{1}{2}II)$. Finally, a player
  $(\tau^{\pi},\overline{\mu})$ maximinimizes the objective fitness game $G^C$ of Figure
  \ref{coordgame1} with flat probability $\overline{\mu}$ over the co-player's actions, i.e.,
  maximizes objective utility given $\overline{\mu}$. In game $G^C$ this yields action $II$.

\begin{figure}
\begin{center}%
\begin{tabular}{|c|c|c|}
\hline 
 & I & II\tabularnewline
\hline 
\hline 
I & 0;0 & -2;-1\tabularnewline
\hline 
II & -1;-2 & 0;0\tabularnewline
\hline 
\end{tabular}\end{center}

\protect\caption{Regret representation $\tau^{reg}(G^C)$}
\label{coordgame1reg}
\end{figure}

We are now in the position of deriving behavior directly from the pair $(\tau, e)$. Definitions \ref{def:utility-belief} and \ref{def:maxminSEU} simply formalize the notions of subjective expected utility and maxmin subjective expective utility, respectively.

\begin{definition} \label{def:utility-belief}

Given a game $G$, an action $a_i \in A_i$, and a belief $\mu \in \lbrace \mu' \in \mathbb{R}^{A_{-i}}: \sum_{a_{-i} \in A_{-i}} \mu'(a_{-i}) = 1 \rbrace$, we define $\tau(G)(a_i, \mu):= \sum_{a_{-i} \in A_{-i}} \tau(G)(a_i, a_{-i}) \cdot \mu(a_{-i})$.

\begin{definition} \label{def:maxminSEU}

Given a game $G$ and a player $(\tau, e)$, we define $\text{maxmin}(\tau(G), e):= \lbrace a'_i \in A_i : a_i \in \text{max}_{a_i \in A_i} \text{ min}_{\mu \in e} \text{ } \tau(G)(a_i, \mu) \rbrace$.

\noindent Definition \ref{def:expected fitness} gives the expected evolutionary fitness of player $(\tau, e)$ against player $(\tau', e')$ in game $G$.\footnote{As already specified in the previous example, we assume that, whenever the set $\text{maxmin}(\tau(G), e)$ is not a singleton, the player $(\tau, e)$ will mix with equal probability among the actions in $\text{maxmin}(\tau(G), e)$.}

\begin{definition} \label{def:expected fitness}

Given a game $G \in \mathcal{G}$, a player $(\tau, e)$, and a player $(\tau', e')$, we define $\pi^G((\tau, e),(\tau', e')):= \pi^G(\text{maxmin}(\tau(G), e),\text{maxmin}(\tau'(G), e'))$.

%\todo[inline]{should we explain here, or very soon, what the resulting types really are: regret minimizers, maximin players \dots}

Many other alternatives different from maxmin expected utility could produce behavior: choice rules like minmax, maxmax, and minmin expected utility would similarly generate behavior from a player type and an epistemic type. Here we postulate that players are maximinimizers for two main reasons. First, as already mentioned, among other possible choice rules maxmin is not only intuitively appealing but also corroborated from an empirical point of view. Secondly, our main interest is the evolution of different patterns of subjective conceptualizations of games (player types), and we want to show that the veridical representation (objective type) is not necessarily the fittest and may not be evolutionarily stable. Assuming that all players use the maxmin rule independent of their player type allows us to factor the choice rule out of the selection process and to focus on player types.

Before concluding this section, it is worth stressing that player $(\tau^{\pi}, \Delta(A_{-i}))$ turns out to be the objective maximinimizer, player $(\tau^{reg}, \Delta(A_{-i}))$ is the regret minimizer, player $(\tau^{\pi}, \overline{\mu})$ is the objective expected utility maximizer, and player $(\tau^{reg}, \overline{\mu})$ is the expected regret minimizer. Moreover, the following facts hold.

\begin{fact} \label{fact:maxEU-minReg}

For any probability measure $\mu$, maximization of expected utility and minimization of expected regret are behaviorally equivalent.

\end{fact}

\noindent As we have seen in the previous example, this is not the case in the presence of (radical) uncertainty, in the sense of representation of belief state in terms of a set of probability, as in $ \Delta(A_{-i}) $.

\begin{fact} \label{fact:equivalence2x2}

In the class of $2 \times 2$ symmetric games, players $(\tau^{reg}, \Delta(A_{-i}))$, $(\tau^{\pi}, \overline{\mu})$, and $(\tau^{reg},\overline{\mu})$ are behaviorally equivalent.

\end{fact}

\noindent Two remarks. Firstly, Fact \ref{fact:equivalence2x2} does not hold for $n \times n$
symmetric games in general. Secondly, we treat behaviorally equivalent types as distinct,
because we assume that what is selected by biological evolution is not choice behavior as such,
but the mechanism which generates behavior. This matters in particular if we assume that
mutations affect different components of the behavior generating mechanisms (see below).

\fi

@@

\noindent Let us consider the case of sampling from a finite set of values first. To show
the result we will use the following partition of $\mathcal{G}$.
\begin{enumerate}
\item Coordination games $\mathcal{G}^{C}$: $a>c$ and $d>b$;
\item Anti-coordination games $\mathcal{G}^{A}$: $a<c$ and $d<b$;
\item Strong dominance games $\mathcal{G}^{S}$: aut $(a>c$ and $b>d)$
aut $(a<c$ and $b<d)$;
\item Weak dominance games $\mathcal{G}^{W}$: aut $a=c$ aut $b=d$;
\item Indifferent games $\mathcal{G}^{I}$: $a=c$ and $b=d$.
\end{enumerate}
Before proving the lemmas, it is convenient to fix some notation.
We denote by $G^{C}$ a coordination game in $\mathcal{G}^{C}$ with
payoffs $a^{G^{C}}$, $b^{G^{C}}$, $c^{G^{C}}$, and $d^{G^{C}}$;
similarly, for $G^{A}$, $G^{S}$, $G^{W}$, and $G^{I}$. Let us
call $I_{G^{C}}^{RM}$ the event that a RM player plays action $I$
in the game $G^{C}$; and similarly for action $II$, for player MM,
and for games $G^{A}$, $G^{S}$, $G^{W}$ and $G^{I}$. Moreover,
let us recall the definition of evolutionarily stable strategy and
neutrally stable strategy.

@@

\noindent We now distinguish between two cases: (1) $ a-c = d-b $ and (2) $  a-c \neq d-b $. Notice that $P(I_{G^{C}}^{RM}\cap II_{G^{C}}^{RM}) \neq 0$ if and only if case (1) obtains, and that $a>d$ and (1) imply $II_{G^{C}}^{MM}$. Then, from (1) we have:
